---
#Based off Steven Miller rmarkdown syllabus template
output: 
  pdf_document:
    #citation_package: natbib
    keep_tex: false
    fig_caption: yes
    latex_engine: pdflatex
    template: svm-latex-syllabus.tex
geometry: margin=1in
bibliography: master.bib
csl: apa6.csl

title: "PSYC 259: Principles of Data Science"
author: Prof. John M. Franchak
date: "Winter 2021"

email: "franchak@ucr.edu"
web: "https://github.com/PSYC-259-Data-Science"
officehours: "By Appointment"
office: "https://ucr.zoom.us/my/johnfranchak"
classroom: "https://ucr.zoom.us/j/97985633776"
classhours: "W  1:30-4:20 PM"

fontfamily: mathpazo
fontsize: 11pt
header-includes:
   - \linespread{1.05}
---

```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.path='figs/')
knitr::opts_chunk$set(cache.path='_cache/')
knitr::opts_chunk$set(warning=F, message=F)
# knitr::opts_knit$get(kable.force.latex = TRUE)
```

```{r dates_bib, include=FALSE}
knitr::opts_chunk$set(cache=FALSE, dev='pdf')
wed <- as.Date("2021-01-06")

advdate <- function(obj, adv) {
 tmon <- obj + 7*(adv-1)
 tfri <- obj + 4 + 7*(adv-1)
 tmon <- format(tmon, format="%m/%d")
 tfri <- format(tfri, format="%m/%d")
 zadv <- sprintf("%02d", adv)
 #tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri)
  tmp <- paste("Week ",zadv,sep='',", ", tmon)
 return(tmp)
}

# library(RefManageR)
# # library(knitcitations)
# # library(rcrossref)
# bib <- ReadBib("master.bib")
# myopts <- BibOptions(bib.style = "authoryear", style="latex", first.inits=FALSE, max.names = 20)

```

# Course Description

Most quantitative courses (importantly) focus on the final steps of data analysis---conducting and understanding statistical tests. However, much of the work in data science is taking raw data, often from multiple, incompatible sources, and processing those data into a usable form. This course will emphasize the importance of robust, documented, and automated workflows for processing data to save time, reduce errors, improve reproducibility, and facilitate collaboration among multiple researchers. We will also spend time on data visualization and communication---an important part of creating, checking, and collaborating on data workflows. We will use the R programming language, Github, and Rmarkdown to work through examples, but the focus is on concepts/best practices that can be applied to any software or programming language. The course is open to students who have little programming experience or experience with R. The goal is for students at all levels of programming experience to set goals to improve their data science skills. 

# Course Objectives

The goals of this course are for you to critically analyze and improve your data analysis workflows. Implementing robust, automated procedures for handling data will allow you to:

- Foster open science through increased transparency, reproducibility, and easier data sharing
- Increase the fidelity of your data and analyses by detecting and preventing errors
- Better understand and communicate about your data
- Save time by preventing errors, automating tasks, and reusing code
- Facilitate collaboration with organized and documented workflows

# Course Materials

- The course [Github page](https://github.com/PSYC-259-Data-Science) has links to project files, readings, and the syllabus
- We will work through examples and complete assignments through a course workspace on [RStudio Online](https://rstudio.cloud/spaces/106138/projects). To sign up, follow this [link](https://rstudio.cloud/spaces/106138/join?access_code=W90ZG4T1F5TjqEW7EloGQQcpp%2BKrTmufayP9G5Hj). You will need to create a free user account. We will go over how to access materials on the first day of class. 
- Readings from *R for Data Science* [@WickhamGrolemund2017] are available [online](https://r4ds.had.co.nz/index.html).
- PDFs of other course readings referenced below are available on Github.

# Course Policies

TBD

## Assignments and Grading

## Remote Instruction

\newpage

# Class Schedule

Readings should be completed prior to each class. 

## `r advdate(wed, 1)`: Course Goals
- Goals of the class
- Logistics (RStudio/Github)
- Readings and assignments

## `r advdate(wed, 2)`: Data Workflow
- File organization
- Version control
- Workflow examples [John's data "Show and Tell"]
- *SKILLS:* Basic I/O functions, using Github
- **READING:** Minimizing Mistakes [@RouderHaaf2019]
<!--Data science book I/O, workspace stuff?
Something on Github/Version Control?-->


##  `r advdate(wed, 3)`: Automation
- Copy/paste is evil (automating data pipeline)
- Drop-down menus are evil (automating your analyses)
- Machine-readable formats
- *SKILLS:* Basic scripting, iteration, merging data 

##  `r advdate(wed, 4)`: Data "Show and Tell"
- Present your current workflow
- Identify places for potential errors
- Identify places to increase efficiency/automation

##  `r advdate(wed, 5)`: Data Structure
- Data types 
- Data organization (variables/observations)
- *SKILLS:* Factors/strings, reshaping/aggregating data
- **READING:** Tidy Data [@Wickham2014]

##  `r advdate(wed, 6)`: Error Checking
- Data validation
- Automating visualizations 
- Exploratory data analysis
- *SKILLS:* Logical statements, indexing, basic visualizations

##  `r advdate(wed, 7)`: Encapsulation: Reusing Functions
- Best option: Use existing APIs and packages
- Next best option: Writing your own general-purpose functions
- Not great option: Writing overly-specific functions that you can never use again
- *SKILLS:* Sourcing functions, parameterization

##  `r advdate(wed, 8)`: Reproducible Reports and Documentation
- Code as documentation
- Documentation faciliates collaboration and data sharing
- Reproducible reports/data capsules preserve analyses
- *SKILLS:* Rmarkdown, papaja, package control
- **READING:** Transparency in psychological science [@KleinHardwicke2018]

##  `r advdate(wed, 9)`: Visualization
- Communicating through graphical styles
- Interactive/animated plots for data exploration
- *SKILLS:* ggplot and extensions

##  `r advdate(wed, 10)`: Project Presentations
- How have you changed your workflow?
- What have you learned about your data? 
- What problems are still unresolved?
\newpage

# References
